---
title: "Building the Count matrix from Salmon quantification data"
author: "Andreé Johnsson"
date: "11/02/2019"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.path='0-Figs/'
                      , include = TRUE
                      , eval = TRUE
                      , echo=TRUE
                      , warning=FALSE
                      , fig.align='center')
```

This procedure is derived from 
<https://bioconductor.org/packages/release/bioc/vignettes/tximport/inst/doc/tximport.html/>

```{r}
load(file ="~/Project-Git-Files-aspen-FTL1-growth-cessation/src/R/1.rawDataQA/0-RData_Storage/1-readInSalmon.RData")
```

```{r echo=FALSE}
suppressPackageStartupMessages(library(magrittr))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(readr))
suppressPackageStartupMessages(library(tximport))
suppressPackageStartupMessages(library(pander))
```

Loading the sample information
```{r eval=FALSE}
sample_info <- 
    read_csv("~/Git/UPSCb/projects/aspen-FTL1-growth-cessation/doc/OVE.Nilsson_18_01_sample_info.csv"
             , col_names = TRUE)
```

```{r}
pander(head(sample_info,3))
```

We read in the files extension names recursively searching for the "quant.sf" files in the
4.salmon directory. Here create a named vector pointing to the quantification files.
```{r eval=FALSE}
file.paths <- list.files("~/Data-Files-aspen-FTL1-growth-cessation/4.salmon"
                   , recursive = TRUE
                   , pattern = "quant.sf"
                   , full.names = TRUE)
```

Now we name the file extensions the same as the NGI.ID to know which sample is which.
```{r}
names(file.paths) <- str_extract(file.paths,"P[0-9]{5}_[0-9]{3}")
head(file.paths,2)
```

Now we reorder the read sample_info dataframe to match with the file names order
```{r eval=FALSE}
sample_info <- sample_info[match(names(file.paths),sample_info$NGI.ID),] 
```

Lets read in the tx2gene, this is the refference transcriptome and assosiates a 
transcipt ID to a gene ID in column 1 and 2 respectively. Important to notice is that 
many transcripts with unique IDs may map to the same gene. Hence the Read/Gene ID mapping is ambiguous. 

```{r eval=FALSE}
tx2gene <- 
    read_tsv("/mnt/picea/storage/reference/Populus-tremula/v1.1/indices/kallisto/tx2gene.tsv"
               , col_names = c("Transcipt.ID","Gene.ID"))
              
```
```{r}
pander(head(tx2gene,3))
```

Lets import the transcript level estimates, by specifying the txt2gene file the type of method "salmon",
we are able to quantify the reads. The intresting thing is that after reading in the quantified data we
obtain the meassage: "transcripts missing from tx2gene: 883".

```{r eval=FALSE}
txi.salmon <- tximport(files = file.paths
                       , type = "salmon"
                       , txOut = TRUE)
```

This is then the count matrix, which tells us how many reads per gene in every sample we have.
We also round the count metrix to integers.
```{r eval=FALSE}
kg.salmon <- round(summarizeToGene(txi.salmon, tx2gene)$counts)
```

```{r}
pander(head(kg.salmon[1:5,1:3]))
```

This procedure produces a count matrix which is the same as with the prevously provided sprict BiologicalQA.R but in a more comprehensive way. (I have checked is the same by all.equal() function)


Export the raw data to CSV
```{r eval=FALSE}
dir.create(file.path("analysis","salmon"),showWarnings = FALSE, recursive = TRUE)
write.csv(kg,file = "analysis/salmon/raw-unnormalised-gene-expression_data.csv")
```


```{r}
kg.salmon %>% as.data.frame() %>% rownames_to_column(var="genes") %>% 
    write_tsv("~/Project-Git-Files-aspen-FTL1-growth-cessation/src/R/restart_Analysis_To_make_int_Understandable/0-Important-DataTables/kg.salmon.tsv", col_names = TRUE)
```


```{r eval=FALSE}
save.image(file ="~/Project-Git-Files-aspen-FTL1-growth-cessation/src/R/1.rawDataQA/0-RData_Storage/1-readInSalmon.RData")
```

```{r}
sessionInfo()
```

