---
title: "Extracting the \"alfredo\" dataset and enlarging the metadata from sample_info"
author: "Andreé Johnsson"
date: "12/02/2019"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.path='Figs/'
                      , include = TRUE
                      , eval = TRUE
                      , echo=TRUE
                      , warning=FALSE
                      , fig.align='center')
```

```{r}
load(file ="~/Project-Git-Files-aspen-FTL1-growth-cessation/src/R/1.rawDataQA/0-RData_Storage/1-readInSalmon.RData")
```


```{r echo=FALSE}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(magrittr))
suppressPackageStartupMessages(library(pander))
```

```{r}
str(sample_info)
```

# Filtering for the "alfredo" dataframe and enlarging the matadata 

We want to take a look at the information for only the "alfredo" data set.
```{r eval=FALSE}
sample_info.alfredo <- filter(sample_info,sample_info$Dataset.ID=="alfredo")
```

Now lets expand the User.ID columns to get some more information of the growth conditions
```{r eval=FALSE}
sample_info.alfredo<- separate(sample_info.alfredo, col = User.ID
                                   , into=c("Genotype","Treatment")
                                   , sep=" ",fill = "right", remove = FALSE)

sample_info.alfredo <- separate(sample_info.alfredo, col = Treatment
                                    , into=c("Treatment","Replicate.ID")
                                    , sep="-",fill = "right", remove = TRUE)

```

```{r}
#str(sample_info.alfredo)
```

Now we make the Genotype and Treatment to be factors and also relevel the Genotype to have 
as default T89.

```{r eval=FALSE}
sample_info.alfredo$Genotype %<>%
    as.factor()

sample_info.alfredo$Genotype <- relevel(sample_info.alfredo$Genotype,"T89")

sample_info.alfredo$Treatment %<>%
    as.factor()
```

```{r}
str(sample_info.alfredo)
```

We give the missing replicate ID the ID of zero
```{r}
sample_info.alfredo$Replicate.ID %<>%
    ifelse(is.na(.),0,.)

#ifelse(is.na(sample_info.alfredo.noBad$Replicate.ID),0,sample_info.alfredo.noBad$Replicate.ID)
```


Setting an order for the tratment factors reprecenting the growing seasson 
This will maybe be done in the future but for now we stick to Treatment just being a factor.
```{r}
#sample_info.alfredo$Treatment %<>% 
#    factor(. , levels= c("LD", "SDW1", "SDW2" ,"SDW3","SDW10", "CTW10", "BB", "LD2")
#                           , ordered = TRUE)

#factor(sample_info.alfredo$Treatment)[1]
```

Now we get rid of the real bad samples leaving on only the "good" and "intermediate" samples
```{r eval=FALSE}
sample_info.alfredo.noBad <- filter(sample_info.alfredo,sample_info.alfredo$Status!="bad")
unique(sample_info.alfredo.noBad$Status)
```
```{r eval=FALSE}
droplevels(sample_info.alfredo.noBad$Status)
```

# How many replicates do we have left? ( -3 *T89 LD2* and an excess of *T89 LD*)

Now we take a look at how many replicates we have for each treatment and phenotype
Comparing the two tables we see that we have removed 3 samples from the *T89 LD2*.
and also that we have an execs of T89 LD. 

## All samples from the "alfredo" dataset
```{r}
pander(table(sample_info.alfredo[,c("Genotype","Treatment")])) 
#pander(table(sample_info.alfredo.noBad[,2:3]))
```

## Only the "good" and "intermediate" samples from the "alfredo" dataset
```{r}
pander(table(sample_info.alfredo.noBad[,c("Genotype","Treatment")]))
```

Saving the separated data
```{r eval=FALSE}
write.csv(sample_info.alfredo
          , file = "~/Project-Git-Files-aspen-FTL1-growth-cessation/doc/sample_info_alfredo.csv"
          , row.names = FALSE)

write.csv(sample_info.alfredo.noBad
          , file = "~/Project-Git-Files-aspen-FTL1-growth-cessation/doc/sample_info_alfredo_noBad.csv"
          , row.names = FALSE)
```

```{r eval=FALSE}
save.image(file =
               "~/Project-Git-Files-aspen-FTL1-growth-cessation/src/R/1.rawDataQA/0-RData_Storage/1-readInSalmon.RData")
```

Why do we have 9 replicates? could i be separate batches?
```{r}
T89.LD.ID <- sample_info.alfredo.noBad %>% filter(Genotype =="T89",Treatment=="LD") %>% select(NGI.ID)
T89.LD.ID$NGI.ID %>% 
    paste(collapse= "|") %>% 
    grep(file.paths, value = TRUE) %>%
    str_remove("/mnt/picea/home/ajohnsson/Data-Files-aspen-FTL1-growth-cessation/4.salmon/") %>%
    str_remove("/quant.sf") %>%
    as.data.frame() %>%
    mutate(ID = rep("T89-LD",9)) %>%
    rename("File_Paths"=".") %>%
    write_tsv("~/Project-Git-Files-aspen-FTL1-growth-cessation/src/R/2.Differential_Expression_of_the_alfredoDataset/9_replicates_file_names.tsv")

```



## Finnaly the sample_info.alfredo.noBad looks like this:
```{r}
as_tibble(sample_info.alfredo.noBad)
```

```{r}
sessionInfo()
```

