---
title: "Assessing Gene Raw Count Distribution"
author: "Andreé Johnsson"
date: "11/02/2019"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.path='Figs/'
                      , include = TRUE
                      , eval = TRUE
                      , echo=TRUE
                      , warning=FALSE
                      , fig.align='center')
```

```{r}
load(file ="~/Project-Git-Files-aspen-FTL1-growth-cessation/src/R/1.rawDataQA/0-RData_Storage/1-readInSalmon.RData")
load(file="~/Project-Git-Files-aspen-FTL1-growth-cessation/src/R/1.rawDataQA/0-RData_Storage/4-RawCountDistribution.RData")
```


```{r echo=FALSE}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(pander))
```

# QC analysis of the Raw Data Distribution

The data is arranged in sample as columns and genes as rows. 
```{r}
pander(head(kg.salmon[1:5,1:3]))
```


## Check how many genes are never expressed
```{r}
sel <- rowSums(kg.salmon) == 0
sprintf("%s%% percent (%s) of %s genes are not expressed",
        round(sum(sel) * 100/ nrow(kg.salmon),digits = 1),
        sum(sel),
        nrow(kg.salmon))
```


## Gene Mean Raw Count Distribution

```{r eval=FALSE}
totalOriginalGeneMean <- apply(kg.salmon,1,function(x) log10(mean(x)))
```

```{r meanRawCountDist, fig.height=4,fig.width=6}
ggplot(as.data.frame(totalOriginalGeneMean), aes(x = totalOriginalGeneMean)) +
    geom_density(alpha=0.4,fill="darkgreen") +
    labs(title = "Gene mean raw counts distribution"
         , x = "Mean raw counts (log10)") + 
    coord_cartesian(ylim = c(0, 0.5)) 
```


## Sample Gene Mean Raw Count Distribution

The same is done for the individual samples
```{r eval=FALSE}
stacked_kg.salmon <- as.data.frame(stack(kg.salmon))
colnames(stacked_kg.salmon) <- c("GeneID","NGI.ID","value")
stacked_kg.salmon$value_log10 <- log10(stacked_kg.salmon$value)
```

```{r}
pander(head(stacked_kg.salmon,3))
```


```{r eval=FALSE}
# We decide the bad smaples by trail and error when plotting all
# the samples at the same time
lessGoodSamples = sort(apply(kg.salmon, 2, mean))[1:9]

# We subset all of the less good samples
stacked_kg.salmon_NotNormal = stacked_kg.salmon[stacked_kg.salmon[,"NGI.ID"] %in% names(lessGoodSamples),]
```


```{r eval=FALSE}
# The label to be used for all plots
densityPlotlabel <- labs(title = "Sample raw counts distribution", x = "Per gene raw counts (log10)")

# Sample Raw Count Distribution
# All Samples at the same time
allsamp <-
    ggplot(stacked_kg.salmon, aes(x=value_log10,col=NGI.ID)) +
    geom_density() +
    guides(col=FALSE) + 
    densityPlotlabel +
    coord_cartesian(xlim = c(0, 4.3), ylim = c(0,0.65)) 

# Only the bad samples
badsamp <-
    ggplot(stacked_kg.salmon_NotNormal, aes(x=value_log10,col=NGI.ID)) +
    geom_density() +
    densityPlotlabel +
    coord_cartesian(xlim = c(0, 4.3), ylim = c(0,0.65)) 

# After removing the bad samples
nobadsamp <-
    ggplot(stacked_kg.salmon[!(stacked_kg.salmon$NGI.ID %in% stacked_kg.salmon_NotNormal$NGI.ID),]
           , aes(x=value_log10,col=NGI.ID)) +
    geom_density() +
    densityPlotlabel +
    guides(col=FALSE) +
    coord_cartesian(xlim = c(0, 4.3), ylim = c(0,0.65)) 

# Table showing the bad samples status.
badSampleTable <-
    sample_info[sample_info$Status!="good",-c(4,6)]

```

```{r sampleRawCountDistribution, fig.height=4, fig.width=6}
#grid.arrange(allsamp, badsamp, nobadsamp, tableGrob(badSampleTable), nrow = 2)
allsamp
badsamp
nobadsamp
```
NOTE: Most samples show similar trend. Some are shifted to the left, because they were not sequenced so deeply.

By the shift in the normal distribution of respective sample, the samples deviating remarkably are classed as "bad" or "intermediate" the samples following the normal distribution are classed as "good".

## The Diviating Samples to be extra carefull with:
```{r}
pander(badSampleTable)
```

```{r eval=FALSE}
save.image("~/Project-Git-Files-aspen-FTL1-growth-cessation/src/R/1.rawDataQA/0-RData_Storage/4-RawCountDistribution.RData")
```


```{r}
sessionInfo()
```

