---
title: "Making a dds dataframe and running Desq2"
author: "Andreé Johnsson"
date: "02/04/2019"
output: html_document
---

# Loading libraries and sample information
```{r}
suppressPackageStartupMessages(library(magrittr))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(DESeq2))
suppressPackageStartupMessages(library(gridExtra))
```

# We have saved the count matrix as tsv 
```{r eval=FALSE}
kg.salmon <- read_tsv("~/Project-Git-Files-aspen-FTL1-growth-cessation/src/R/restart_Analysis_To_make_int_Understandable/0-Important-DataTables/kg.salmon.tsv", col_names = TRUE)
kg.salmon %<>% column_to_rownames(var="genes")
```
# We have saved the sample dataframe as csv 
```{r eval=FALSE}
sample_info_alfredo_noBad_noBB <- read_csv("~/Project-Git-Files-aspen-FTL1-growth-cessation/src/R/restart_Analysis_To_make_int_Understandable/0-Important-DataTables/sample_info_alfredo_noBad_noBB.csv", col_names = TRUE)
```

# Relevel the important genotypes and timepoints
```{r eval=FALSE}
sample_info_alfredo_noBad_noBB %<>%
    transmute(NGI.ID,
           Genotype = factor(Genotype,levels = c("T89","SVP","FT")) ,
           TimePoint = factor(TimePoint, levels= c("LD", "SDW1", "SDW2", "SDW3", "SDW10", "CTW10", "LD2")),
           GT = factor(GT,levels = c("T89-LD", "T89-SDW1", "T89-SDW2","T89-SDW3","T89-SDW10"
                                     ,"T89-CTW10","T89-LD2", "SVP-LD", "SVP-SDW1", "SVP-SDW2"
                                     ,"SVP-SDW3", "SVP-SDW10", "SVP-CTW10" ,"SVP-LD2","FT-LD"
                                     ,"FT-SDW1", "FT-SDW2", "FT-SDW3","FT-SDW10","FT-CTW10","FT-LD2") ) )

sample_info_alfredo_noBad_noBB %>% map(~levels(.x))
```

# Subsetting the countMatrix for all samples to only include "Afredo" dataset
```{r eval=FALSE}
kg_alfredo_noBad_noBB <- kg.salmon %>% select(sample_info_alfredo_noBad_noBB$NGI.ID)
```

# Reordering the rows in sample_info to match with the columns in the countMatrix
```{r eval=FALSE}
sample_info_alfredo_noBad_noBB <-
    sample_info_alfredo_noBad_noBB[
        match(sample_info_alfredo_noBad_noBB$NGI.ID,colnames(kg_alfredo_noBad_noBB)),]
```

# Creating a dds matrix with all the relevant information in with the right levels
```{r eval=FALSE}
dds_alfredo_noBad_noBB <- DESeqDataSetFromMatrix(
    countData = kg_alfredo_noBad_noBB,
    colData = sample_info_alfredo_noBad_noBB,
    design = ~ Genotype + TimePoint + Genotype:TimePoint )
``` 
# Now we run the D.E
```{r eval=FALSE}
dds_alfredo_noBad_noBB %<>%  DESeq()
```

```{r}
resultsNames(dds_alfredo_noBad_noBB)
```

# Variance Stabilising Transformation with respect to the model
```{r eval=FALSE}
vsd_alfredo_noBad_noBB_aware <- varianceStabilizingTransformation(dds_alfredo_noBad_noBB, blind = FALSE)
vst_alfredo_noBad_noBB_aware <- assay(vsd_alfredo_noBad_noBB_aware)
vst_alfredo_noBad_noBB_aware <- vst_alfredo_noBad_noBB_aware - min(vst_alfredo_noBad_noBB_aware)
```



```{r eval=FALSE}
vst_alfredo_noBad_noBB_aware %>% as.data.frame() %>% rownames_to_column(var="genes") %>%
    write_csv("~/Project-Git-Files-aspen-FTL1-growth-cessation/src/R/restart_Analysis_To_make_int_Understandable/0-Important-DataTables/vst_alfredo_noBad_noBB_aware.csv", col_names= TRUE)
```


```{r}
resultsNames(dds_alfredo_noBad_noBB)
```

