---
title: "Extract_the_sig_genes_matrixs"
author: "Andreé Johnsson"
date: "02/04/2019"
output: html_document
---

# Loading libraries and sample information
```{r}
suppressPackageStartupMessages(library(magrittr))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(DESeq2))
suppressPackageStartupMessages(library(gridExtra))
```

```{r}
load("~/Project-Git-Files-aspen-FTL1-growth-cessation/src/R/restart_Analysis_To_make_int_Understandable/dds_alfredo_noBad_noBB.RData")
```


```{r}
resultsNames(dds_alfredo_noBad_noBB)
```

# The relevant combinations we want, between sucessive timepoints 
```{r}
comb <- c(
"GenotypeSVP.TimePointSDW1-Genotype_SVP_vs_T89", # This interaction represents LD-SDW1 in SVP
"GenotypeSVP.TimePointSDW2-GenotypeSVP.TimePointSDW1", 
"GenotypeSVP.TimePointSDW3-GenotypeSVP.TimePointSDW2", 
"GenotypeSVP.TimePointSDW10-GenotypeSVP.TimePointSDW3", 
"GenotypeSVP.TimePointCTW10-GenotypeSVP.TimePointSDW10",
"GenotypeSVP.TimePointLD2-GenotypeSVP.TimePointCTW10",

"GenotypeFT.TimePointSDW1-Genotype_FT_vs_T89", # This interaction represents LD-SDW1 in FT
"GenotypeFT.TimePointSDW2-GenotypeFT.TimePointSDW1", 
"GenotypeFT.TimePointSDW3-GenotypeFT.TimePointSDW2", 
"GenotypeFT.TimePointSDW10-GenotypeFT.TimePointSDW3", 
"GenotypeFT.TimePointCTW10-GenotypeFT.TimePointSDW10",
"GenotypeFT.TimePointLD2-GenotypeFT.TimePointCTW10",

"Genotype_SVP_vs_T89-Genotype_FT_vs_T89", # This interaction represents the difference in LD
"GenotypeSVP.TimePointSDW1-GenotypeFT.TimePointSDW1",
"GenotypeSVP.TimePointSDW2-GenotypeFT.TimePointSDW2",
"GenotypeSVP.TimePointSDW3-GenotypeFT.TimePointSDW3",
"GenotypeSVP.TimePointSDW10-GenotypeFT.TimePointSDW10", 
"GenotypeSVP.TimePointCTW10-GenotypeFT.TimePointCTW10",
"GenotypeSVP.TimePointLD2-GenotypeFT.TimePointLD2"
)
```


```{r}
get_Two_M.2 <- function(dds, comb){
    colnames = NULL
    # Make two matrix of NA with gene as rownames
    log2f.matrix <- (assay(dds)[,1:length(comb)] * NA) 
    colnames(log2f.matrix) = comb
    padj.matrix <- log2f.matrix
    
    # Gets every interaction and fills the NA matrix accordinly with the real
    # values 
    for (i in comb){
        res = results(object = dds
                  , contrast = strsplit(i,"-")
                  , alpha = 0.01
                  , lfcThreshold = 0.5)
        
        log2f.matrix[,i] = res[,"log2FoldChange"]
        padj.matrix[,i] = res[,"padj"] 
        
        # Trims the colnames 
        colname_tmp = i %>% gsub("Genotype|TimePoint","",.) %>%
            str_split("\\.|-") %>% map_chr(~unique(.x) %>% paste(collapse ="_"))
        colnames %<>% append(colname_tmp)
        
        print(colname_tmp)
    }  
    
    colnames(log2f.matrix) = colnames
    colnames(padj.matrix) = colnames
    # Returns the two matrix of all given combination with
    # respective log2foldchange values and padj values, this is later
    # trimed to obtained the significant genes.
    return( list( "log2FC" = log2f.matrix,"padj" = padj.matrix) )
}
```

```{r}
trimlog2foldMatrix <- function(test){
    # Trims Log2fold matrix to only include the significantgenes with their
    # respective log2foldchange value, the rest is NA.
    log2f.matrix = test$log2FC
    padj.matrix = test$padj
    
    logical_matrix = (padj.matrix<= 0.01 &
                                !is.na(padj.matrix) &
                                abs(log2f.matrix) >= 0.5 )
    for (col in 1:ncol(log2f.matrix)){
        
        log2f.matrix[ logical_matrix[,col] == FALSE , col] = NA
        }
    return(as.data.frame(log2f.matrix))
    }
```


# First we get the two matrices of Log2FC and padj for all the desire combinations
```{r}
log2FC_and_padj_matrices = get_Two_M.2(dds_alfredo_noBad_noBB,comb = comb)
```

# We change the name of _SVP_vs_T89__FT_vs_T89 to SVP_LD_FT since it reprecents that interacation.
```{r}
log2FC_and_padj_matrices$log2FC %<>%
    as.data.frame() %>%
    dplyr::rename(SVP_LD_FT= `_SVP_vs_T89__FT_vs_T89`
                  , SVP_SDW1_LD = `SVP_SDW1__SVP_vs_T89`
                  , FT_SDW1_LD = `FT_SDW1__FT_vs_T89`)
log2FC_and_padj_matrices$padj %<>%
    as.data.frame() %>%
    dplyr::rename(SVP_LD_FT= `_SVP_vs_T89__FT_vs_T89`
                  , SVP_SDW1_LD = `SVP_SDW1__SVP_vs_T89`
                  , FT_SDW1_LD = `FT_SDW1__FT_vs_T89`)
```

# Then we trimm the Log2FC matrix to only include the significant genes
```{r}
trimmed_log2FC_only_sig_genes_matrix <- trimlog2foldMatrix(log2FC_and_padj_matrices)
```


```{r}
all.equal(colnames(log2FC_and_padj_matrices$padj),colnames(log2FC_and_padj_matrices$log2FC),colnames(trimmed_log2FC_only_sig_genes_matrix))
```

# Now we save it in an R object
```{r}
# save(log2FC_and_padj_matrices,trimmed_log2FC_only_sig_genes_matrix,file= "~/Project-Git-Files-aspen-FTL1-growth-cessation/src/R/restart_Analysis_To_make_int_Understandable/log2FC_and_padj_matrices_and_trimmed.RData")
```

# Now we save it in an csv
```{r}
trimmed_log2FC_only_sig_genes_matrix %>% rownames_to_column(var="genes") %>%
    write_csv("~/Project-Git-Files-aspen-FTL1-growth-cessation/src/R/restart_Analysis_To_make_int_Understandable/0-Important-DataTables/trimmed_log2FC_only_sig_genes_matrix.csv", col_names = TRUE)
```

