---
title: "3-plotting-individual-genes"
author: "Andreé Johnsson"
date: "02/04/2019"
output: html_document
---

# Loading libraries and sample information
```{r}
suppressPackageStartupMessages(library(magrittr))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(DESeq2))
suppressPackageStartupMessages(library(gridExtra))
```

# Loading from RData to skip refactoring
```{r}
load("~/Project-Git-Files-aspen-FTL1-growth-cessation/src/R/restart_Analysis_To_make_int_Understandable/sample_info_alfredo_noBad_noBB.RData")
```

# Loading the main table to seach for genes
```{r}
pop_arab_genie_all_genes <- read_csv2("~/Project-Git-Files-aspen-FTL1-growth-cessation/src/R/restart_Analysis_To_make_int_Understandable/0-Important-DataTables/pop_arab_genie_all_genes.csv", col_names = TRUE, guess_max = 80000)
```
# Loading from dataframe
```{r}
median_vst_alfredo_noBad_noBB_aware <-
    read_csv("~/Project-Git-Files-aspen-FTL1-growth-cessation/src/R/restart_Analysis_To_make_int_Understandable/0-Important-DataTables/median_vst_alfredo_noBad_noBB_aware.csv", col_names = TRUE)

vst_alfredo_noBad_noBB_aware <-
    read_csv("~/Project-Git-Files-aspen-FTL1-growth-cessation/src/R/restart_Analysis_To_make_int_Understandable/0-Important-DataTables/vst_alfredo_noBad_noBB_aware.csv", col_names = TRUE)
```

# Rearranging the data to be easily plottable
```{r}
vst_to_plot_all_genes <-
    vst_alfredo_noBad_noBB_aware %>% column_to_rownames(var="genes") %>% t() %>% as.data.frame() %>% 
    rownames_to_column(var="NGI.ID") %>% right_join(sample_info_alfredo_noBad_noBB,.,by = "NGI.ID") 
```

# Relevel the median df to have the correct levels
```{r}
median_vst_alfredo_noBad_noBB_aware %<>%
    mutate(
        Genotype = factor(Genotype, levels = c("T89","SVP","FT")) ,
        TimePoint = factor(TimePoint, levels= c("LD", "SDW1", "SDW2", "SDW3", "SDW10", "CTW10", "LD2"))
    )
```

# Check the levels to get nice plots
```{r}
median_vst_alfredo_noBad_noBB_aware[,1:2] %>% map(~levels(.x))
vst_to_plot_all_genes[,2:3] %>% map(~levels(.x)) 
```


# Plot function for individual genes
```{r}
plotIndividualExpression_with_scatter<- function(vst,median,gene,scatter= FALSE){
    # Both graphs share the same aestetics
    aes.1 = aes_string(x= "TimePoint", y= gene
                          , group= "Genotype", col= "Genotype")
    # Plots the median of the replicates across time
    gene_plot <-
        ggplot(median,aes.1) +
        facet_wrap(~Genotype, scales= "free_y") +
        geom_point(size=2) +
        geom_line(size=1.1) +
        theme_bw() +
        theme(axis.text.x = element_text(angle = 50, hjust = 1) ) +
        scale_colour_brewer(palette= "Dark2")  +
        labs(title= "Expression of The Median Of The Replicates")
    # Plots the replicatres themselfs if scatter is true
    if(scatter == TRUE){
        gene_plot = gene_plot + geom_point(data = vst, aes.1)
        return(gene_plot)
        }
    return(gene_plot)
}
```

# Search for a gene in the big table
```{r}
SVP_genes_ids <- pop_arab_genie_all_genes %>% filter(str_detect(All.Gene.Simbol.tair,"SHORT VEGETATIVE PHASE")) %>% pull(Gene) %>% unique()

FT_genes_ids <- pop_arab_genie_all_genes %>% filter(str_detect(All.Gene.Simbol.tair,"FLOWERING LOCUS T")) %>% pull(Gene) %>% unique()
```

# Plot a list of genes in the consol
```{r}
lapply(SVP_genes_ids, function(x){
plotIndividualExpression_with_scatter(vst = vst_to_plot_all_genes
                                      , median = median_vst_alfredo_noBad_noBB_aware
                                      , gene= x
                                      , scatter = FALSE) })

```


# Save png, saves the png in the right format and size
```{r}
save_png_of_genes <- function(path, genes, title="",scatter=FALSE){
    no.genes = length(genes)
    grobs <- lapply(genes, function(x){
        plotIndividualExpression_with_scatter(vst = vst_to_plot_all_genes
                                      , median = median_vst_alfredo_noBad_noBB_aware
                                      , gene= x
                                      , scatter = scatter) })

    png(path, height = (280*no.genes), width = 900, units = "px", res = 100)
    grid.arrange(grobs=grobs, ncol=1,
      top = textGrob(title, gp=gpar(fontsize=16,font=3)) )
    dev.off()   

}
```




```{r}
save_png_of_genes("SVP_genes_ids.png",SVP_genes_ids,"SVP Genes")
save_png_of_genes("FT_genes_ids.png",FT_genes_ids,"FT Genes")
```


