---
title: "7.1 Preprocecing data to perform soft Clustering"
author: "Andreé Johnsson"
date: "11/04/2019"
output: html_document

---

<https://2-bitbio.com/post/clustering-rnaseq-data-using-fuzzy-c-means-clustering//>

# Loading libraries and sample information
```{r}
suppressPackageStartupMessages(library(Mfuzz))
suppressPackageStartupMessages(library(magrittr))
suppressPackageStartupMessages(library(tidyverse))
```


```{r}
median_vst_alfredo_noBad_noBB_aware <-
    read_csv("0-Important-DataTables/median_vst_alfredo_noBad_noBB_aware.csv", col_names = TRUE)
```

# Make nice row names and get rid of unnececarry columns
```{r}
median_vst_alfredo_noBad_noBB_aware %<>%
    mutate(GT = map2_chr(Genotype, TimePoint, ~paste(.x,.y,sep=" ")) ) %>%
    column_to_rownames(var="GT") %>%
    select(-Genotype,-TimePoint)
```

# Logic vector of genes with non zero variance
```{r}
var.per.gene.sorted <-
    median_vst_alfredo_noBad_noBB_aware %>%
    map_dbl(~var(.x)) %>% 
    sort(decreasing = TRUE) %>%
    .[.>0]

summary(var.per.gene.sorted)
sum(var.per.gene.sorted > 0.352606) 
```

# Filter away the non variant genes
```{r}
median_vst_ordered_by_var <- median_vst_alfredo_noBad_noBB_aware[,names(var.per.gene.sorted)]
dim(median_vst_ordered_by_var)
```


# Standard Deviation of gene expressoin vectors before standarization
```{r}
median_vst_ordered_by_var %>%
    map_dbl(~sd(.x)) %>%
    base::sort(decreasing = FALSE) %>%
    plot(1:length(.),.,xlab="Ordered Genes",ylab="Sd") 
```

# Standarize across genes
```{r}
median_vst_ordered_by_var_filtered_scaled <- scale(median_vst_ordered_by_var)
median_vst_ordered_by_var_filtered_scaled %<>% t() %>% as.matrix() 
dim(median_vst_ordered_by_var_filtered_scaled)
```

# Here we see the effect of scaling and filtering

The distribution of the gene expression is not longer skewed
it is balanced.
```{r}
quantile(as.matrix(t(median_vst_ordered_by_var)))
quantile(median_vst_ordered_by_var_filtered_scaled)
```

```{r}
median_vst_ordered_by_var_filtered_scaled %>%
    as.data.frame() %>% 
    rownames_to_column("genes") %>%
    write_tsv("0-Important-DataTables/Mfuzz_median_vst_ordered_by_var_filtered_scaled.tsv"
              , col_names = TRUE)

```


