---
title: "Gene enrichment"
author: "Andreé Johnsson"
date: "28/02/2019"
output: html_document
---

<http://revigo.irb.hr/revigo.jsp/>

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.path='0-Figs/'
                      , include = TRUE
                      , eval = TRUE
                      , echo=TRUE
                      , warning=FALSE
                      , fig.align='center')
```

Loading packages and data
```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(magrittr))
suppressPackageStartupMessages(library(parallel))
suppressPackageStartupMessages(library(dendextend))
suppressPackageStartupMessages(library(KEGGREST))
suppressPackageStartupMessages(library(gplots))
suppressPackageStartupMessages(library(hyperSpec))
suppressPackageStartupMessages(library(LSD))
suppressPackageStartupMessages(library(matrixStats))
suppressPackageStartupMessages(library(pander))
suppressPackageStartupMessages(library(RColorBrewer))
suppressPackageStartupMessages(library(wordcloud))
```



```{r}
load("~/Project-Git-Files-aspen-FTL1-growth-cessation/src/R/2.Differential_Expression_of_the_alfredoDataset/0-RData_Storage/5-gopher.RData")

suppressPackageStartupMessages(source("~delhomme/Git/UPSCb/src/R/gopher.R"))
```

Now we select the background genes, in this case by selecting with feature selection 
instead of rowSums>0 we hopefully get a more acurate result. We get at least 7948 lesser genes selected by feature selection. 

```{r}
bg.pop <- kg.salmon[rowSums(kg.salmon)>0,] %>% rownames()
#bg.pop <- kg.salmon[feature.sel,] %>% rownames()
length(bg.pop)
```

```{r}
length(sig.genes.time.interaction)
```


```{r eval=FALSE}
system.time(
    gopher.sig.genes.time.interaction <- 
        map(sig.genes.time.interaction
            , ~ gopher(genes = .x
                       , task = list("go","kegg","pfam")
                       , background = bg.pop
                       , url="potra"))
    )
```

Extrakting the GO term IDs and padj for the interaction terms and put it all together on
a nested data frame. This also filters out the terms dependent on the enrichment,"go","kegg or"pfam" with 
respective significant values of, 0.05, 0.1, 0.01.

```{r}
Get.gopherIDs.2 <- function(gopher.list,term="go"){
    # extrakts the selected go, pfam, kegg data frames from each interaction and 
    # specifically the go id and padj
    # then the numbers of each interacton to make a belonging vector
    # sums all go terms for all interactions in one singel nested data frame
    if (term =="pfam"){
        sel = c("id","name","padj")}
    else{
        sel= c("id","padj")}
    
    df.go.padj <- gopher.list %>% map(term) %>% map_dfr(extract, sel)
    nb.go <- gopher.list %>% map(term) %>% map(nrow) %>% unlist()
    
    condition.vec = NULL
    for (i in 1:length(nb.go)){
        tmp = rep(names(nb.go)[i],nb.go[i])
        condition.vec %<>%  append(tmp)
    }    
    df.go.padj %<>%
        mutate(condition = condition.vec) %>%
        filter(padj<ifelse(term=="go",0.05,ifelse(term=="kegg",0.1,ifelse(term=="pfam",0.01,NULL))) )
    
    df.go.padj %<>% group_by(condition) %>% nest()
    df.go.padj %<>% mutate(n = unlist(map(data,~nrow(.)))) %>% arrange(desc(n))
    return(df.go.padj)
}
```

```{r}
savegopher.files <- function(nest,dir){
    # supplying a nested fata frame and a direcotry so dave all files.
    apply(nest,1, function(row){
        write_tsv(as.data.frame(row[2])
                  , paste0("~/Project-Git-Files-aspen-FTL1-growth-cessation/src/R/2.Differential_Expression_of_the_alfredoDataset/",dir, row[1],".tsv")
                  , col_names = FALSE)})
}
```

Extracting the terms with significant p-values.
```{r}
go.terms = Get.gopherIDs.2(gopher.sig.genes.time.interaction,"go")
kegg.terms = Get.gopherIDs.2(gopher.sig.genes.time.interaction,"kegg")
pfam.terms = Get.gopherIDs.2(gopher.sig.genes.time.interaction,"pfam")
```

Saving files to directory.
```{r eval=FALSE}
savegopher.files(go.terms,"0-go-term-files/")
savegopher.files(pfam.terms,"0-pfam-term-files/")
savegopher.files(kegg.terms,"0-kegg-term-files/")
```


```{r}
plot.gopher.no <- function(nest){
    ggplot(nest,aes(x=fct_reorder(condition,n),y=n)) +
    geom_col() +
    coord_flip() +
    labs(title = "Frequency of terms per condition"
         , y = "No. of terms"
         , x = "Condition")}
```


```{r}
map(list(go=go.terms,pfam=pfam.terms,kegg=kegg.terms),plot.gopher.no)
```

# Kegg

Get all of the ids per time point from the nested list and 
give names to the result to keep time points.

```{r eval=FALSE}
kegg.ids <- kegg.terms$data %>% map(extract,"id")
names(kegg.ids) <- kegg.terms$condition
```

Get all the uniq pathways ids
```{r eval=FALSE}
kegg.pathways <- map(kegg.ids,~ keggLink("pathway",.$id) %>% unique())
```

Now for each time point we want to get the pathway of the the corresponding id and then
extrakt the ENTRY and NAME forming a tibble. Then we wan to count the number of occurances if
the pathway NAME and scale it to 1.

```{r eval=FALSE}
wordc.mapper = as_mapper(~ map_dfr(.x,
                   ~ keggGet(.x) %>%
                       map(extract,c("ENTRY","NAME")) %>%
                       do.call(rbind,.) %>% as_tibble() %>%
                       unnest()) %>%
    count(NAME) %>%
    mutate(n = n/sum(n) ) )
```

This valirable stores the uncounted tables.
```{r eval=FALSE}
wordc.tables.uncounted$`SVP.LD2-SVP.SDW10` %>% head()
```

Now we get a table of Entry and name for each condition
```{r eval=FALSE}
wordc.tables = map(kegg.pathways,wordc.mapper)
names(wordc.tables) <- names(kegg.pathways)
```

```{r}
wordc.tables %>% names()
```


```{r}
# hpal <- ?colorRampPalette(colors = c("blue","white","red"))(10)
# cpal <- rainbow(12)

wordc.tables %>%
    map(~wordcloud(.x$NAME,.x$n, scale = c(1,.01), rot.per = 0))
```

# PFAM

Each pfam,go or kegg is ordered by number of entries in the data cell, by a decending order. That is why
The order or the pictures differ.
```{r}
pfam.names <- pfam.terms$data %>% map(extract,"name") %>% map(~count(.x,name))
names(pfam.names) <- pfam.terms$condition
```

```{r}
pfam.names %>% names()
```


```{r}
pfam.names %>%
    map(~wordcloud(.x$name,.x$n, scale = c(1,.01), rot.per = 0))
```




